---
output: 
  github_document:
    toc: true
    toc_depth: 2
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ResearchGroupTools

[![Build Status](https://travis-ci.org/sfeuerriegel/ResearchGroupTools.svg?branch=master)](https://travis-ci.org/sfeuerriegel/ResearchGroupTools)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ResearchGroupTools)](https://cran.r-project.org/package=ResearchGroupTools)
[![Coverage Status](https://img.shields.io/codecov/c/github/sfeuerriegel/ResearchGroupTools/master.svg)](https://codecov.io/github/sfeuerriegel/ResearchGroupTools?branch=master)

**ResearchGroupTools** provides a collection of utilitiy function for rapid prototyping. These functions facilitate implemenation works related to advanced analytics. As such, it specifically supports data handling, preprocessing, visualization and analytics. 

## Installation

Using the **devtools** package, you can easily install the latest development version of **SentimentAnalysis** with

```{r,eval=FALSE}
install.packages("devtools")

# Option 1: download and install latest version from "GitHub"
devtools::install_github("sfeuerriegel/ResearchGroupTools")

# Option 2: install directly from bundled archive
# devtoos::install_local("ResearchGroupTools_0.1.0.tar.gz")
```

Notes: 

* In the case of option 2, you have to specify the path either to the directory of **ResearchGroupTools** or to the bundled archive **ResearchGroupTools_0.1.0.tar.gz**

* The package will only be shipped via GitHub; CRAN support is not intended. 

## Usage

This section shows the basic functionality of how to perform a sentiment analysis. First, load the corresponding package **ResearchGroupTools**. 

```{r, message=FALSE}
library(ResearchGroupTools)
```

By default, the seed for the random number generator is initialized to 0. 

# Functionality

## Library handling

* `Library()` (note the capital "L") loads packages. If not available, these are automatically installed. 

```{r Library, message=FALSE}
Library("ggplot2", "dplyr")
```


## Numerical functions

* `ceil()` computes the largest integer less or equal given a numerical value. It is a wrapper for `ceiling` with a more consistent naming.

```{r ceil}
ceil(3.4)
```

## Time series

* `differences()` calculates lagged differences of a given order. It is more convenient thant `diff()` as it adds leading `NA` values.

```{r differences}
differences(1:10)
differences(c(1, 2, 4, 8, 16, 32))
differences(c(1, 2, 4, 8, 16, 32), order = 2)
differences(c(1, 2, 4, 8, 16, 32), na_padding = FALSE)
```

* `returns()` calculates returns of a time series (similar to `diff()` for differenes).

```{r returns}
returns(1:10)
returns(c(1, 2, 4, 8, 16, 32))
returns(c(1, 2, 4, 8, 16, 32), na_padding = FALSE) # remove trailing NA's
```

* `logReturns()` computes log-returns (by default, with base `exp(1)`).

```{r log-returns}
logReturns(c(1, 2, 4, 8, 16, 32), base = 2)
```

## Visualization

* `scientific_labels()` enables a nice exponential notation in **ggplot2** plots.

```{r scientific_labels}
library(ggplot2)
df <- data.frame(x=rnorm(100), y=rnorm(100))
ggplot(df, aes(x=x, y=y)) +
  geom_point() +
  scale_x_continuous(labels=scientific_labels) +
  scale_y_continuous(labels=scientific_labels)
```

## Package development

* `remakePackage()` builds, loads and checks package during the development process all at once. In particular, the manual is updated. 

```{r remakePackage, eval=FALSE}
remakePackage()
remakePackage(TRUE) # also runs README.Rmd
```


### License

**ResearchGroupTools** is released under the [MIT License](https://opensource.org/licenses/MIT)

Copyright (c) 2016 Stefan Feuerriegel
