---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ResearchGroupTools

[![Build Status](https://travis-ci.org/sfeuerriegel/ResearchGroupTools.svg?branch=master)](https://travis-ci.org/sfeuerriegel/ResearchGroupTools)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ResearchGroupTools)](https://cran.r-project.org/package=ResearchGroupTools)
[![Coverage Status](https://img.shields.io/codecov/c/github/sfeuerriegel/ResearchGroupTools/master.svg)](https://codecov.io/github/sfeuerriegel/ResearchGroupTools?branch=master)

**ResearchGroupTools** provides a collection of utilitiy function for rapid prototyping. These functions facilitate implemenation works related to advanced analytics. As such, it specifically supports data handling, preprocessing, visualization and analytics. 

## Installation

Using the **devtools** package, you can easily install the latest development version of **SentimentAnalysis** with

```{r,eval=FALSE}
install.packages("devtools")

# Option 1: download and install latest version from "GitHub"
devtools::install_github("sfeuerriegel/ResearchGroupTools")

# Option 2: install directly from bundled archive
# devtoos::install_local("ResearchGroupTools_0.1.0.tar.gz")
```

Notes: 

* In the case of option 2, you have to specify the path either to the directory of **ResearchGroupTools** or to the bundled archive **ResearchGroupTools_0.1.0.tar.gz**

* The package will only be shipped via GitHub; CRAN support is not intended. 

## Usage

This section shows the basic functionality of how to perform a sentiment analysis. First, load the corresponding package **ResearchGroupTools**. 

```{r, message=FALSE}
library(ResearchGroupTools)
```

## Numerical functions

* `ceil` computes the largest integer less or equal given a numerical value. It is a wrapper for `ceiling` with a more consistent naming.

```{r}
ceil(3.4)
```


## Visualization

* `scientific_labels`

```{r scientific_labels}
library(ggplot2)
df <- data.frame(x=rnorm(100), y=rnorm(100))
ggplot(df, aes(x=x, y=y)) +
  geom_point() +
  scale_x_continuous(labels=scientific_labels) +
  scale_y_continuous(labels=scientific_labels)
```


## License

**ResearchGroupTools** is released under the [MIT License](https://opensource.org/licenses/MIT)

Copyright (c) 2016 Stefan Feuerriegel
